\SW: II. New lens candidates from the CFHTLS discovered through citizen science

### 2015-03-22 doc/sw-cfhtls.tex
first draft with all sections, figs, tables and appendix

The CFHT Legacy Survey has been searched for strong lenses with
semi-automated algorithms both at galaxy and groups scales. With the aim
of improving these lens finding robots, we carry out a blind lens search
in the complete \cfhtls \textsc{wide} survey with \sw. We describe the training
sample used for both training the citizen scientists that participated
in \sw and calibrating their performance. We generate realistic looking
simulated samples of lenses both at galaxy and group-scales as part of
this training sample. We present 61 new strong gravitational lens
candidates discovered from the \sw-\cfhtls search out of which 40
candidates are promising. Furthermore, we compile a sample of false
positives which can be used for robot testing, first of its kind for the
lensing community. The simulated sample is also used to make a
comparison between human performance against the robots. We find that
the XX per cent of the known lens sample is recovered and the new lens
sample has XX completeness with respect to the simulated sample. We make
available the following data products: simulated sample, simulation code
and the false positives sample at XXX.

\end{abstract}

@@

The last few decades have seen a rise in the discoveries of strong
gravitational lenses owing to the plethora of interesting applications
they have in astrophysics and cosmology. Strong lenses are routinely
used to probe the dark matter distribution from galaxy (ref) to cluster
scales (ref), to  study distant young galaxies by using the lensing
magnification as a natural telescope (ref), to test the cosmological
model by constraining cosmological parameters such as the Hubble
constant (ref) and dark energy (ref) and many more. Even though strong
lenses are rare, since a foreground massive object needs to be
sufficiently aligned with a distant background source to produce
multiple images, systematic lens searches have led to discovery of over
500 lenses till date (XXX add mld url).

Rarity of lenses implies searching for them is a painstaking task.
Efficient automated methods are thus imperative to finding a reasonably
complete and pure sample of strong lenses.

Since the inception of the first citizen science project, Galaxy Zoo, to
classify galaxy morphologies \citep{Lintott2008}, several astronomy
and non-astronomy projects have been launched by the Zooniverse leading
to many interesting project. For example, PlanetHunters
has discovered XXX transiting exoplanets \citep{Schwamb2012},
Supernova XXX add ref and XXX add ref

In Paper I, we describe \sw, an online system that enables crowd-sourced
detection of gravitational lenses. In this paper (referred to as Paper
II), we describe our first lens search with \sw in the \cfhtls data.

This paper is organised as follows. In \Sref{sec:sw} we give a brief
overview of the \sw system, focusing on the aspects most relevant to the
interpretation of the results of this first lens search. In
\Sref{sec:data} we introduce the \cfhtls imaging data and the known lens
samples from the \cfhtls. In \Sref{sec:}, we explain the training sample
generated to aid the \sw users in the lens search. In
\Sref{sec:results}, we present the new lens candidates and our findings.
We discuss the implications of our results for future lens searches in
\Sref{sec:discuss} and draw conclusions in \Sref{sec:conclude}.

@@

The Canada-France-Hawaii Telescope Legacy Survey (CFHTLS) is a
photometric survey in five optical bands ($u^*g'r'i'z'$) carried out
with the wide-field imager MegaPrime which has a 1~deg$^2$ field-of-view
and a pixel size of 0.186\arcsec. The \cfhtls \textsc{WIDE} covers a total area
of 171~deg$^2$ on the sky and it consists of four fields W1, W2, W3 and
W4. The field W1 has the largest sky coverage of 63.65~deg$^2$. The
fields W2 and W4 have similar sky coverages of 20.32~deg$^2$ and
20.02~deg$^2$, respectively\footnote{These numbers are estimated from http://terapix.iap.fr/cplt/table\_syn\_T0006.html}.
The field W3 has a sky coverage of 42.87~deg$^2$ and is more than twice
as large as W2 and W4.

@@

We note that the \cfhtls is a niche survey with a unique combination of
wide imaging with deep sensitivity. It is a precursor to the ongoing
wide imaging surveys such as the DES, KIDS and HSC and planned surveys
such as the LSST. Searching for lenses with \sw in the \cfhtls will
teach us important lessons and help prepare us for these larger imaging
surveys.

@@

Our aim is to do a blind lens search over the entire \cfhtls
\textsc{wide}. We use the $g$, $r$ and $i$-band color imaging which are
most useful for visual identification of lenses (and provide access to
the $u$ and $z$ band imaging through dashboard with additional tools
XXX).

We make color composite images using publically-available
code\footnote{The open source  color image composition code used in this
work is available from \texttt{http://github.com/drphilmarshall/HumVI}}
following the prescription of \citet{Lupton2004}. Specifically, we first
rescale the pixel values of each channel image into flux units, and
then apply an arcsinh stretch. The stretch parameters are chosen
using a small random sample of images, to ensure that the background
noise is just visible, and that the centres of bright, intermediate
resdhift galaxies are not saturated. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to
allow straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) with an
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts resulting into $\sim$ 430 000 cutouts for the entire \cfhtls
\textsc{wide}. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. It is possible that a lens
candidate happens to be too close to the edge of a cutout, the overlap
between neighbouring cutout allows the inspector to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, an inspector may not necessarily come
across the neighbouring cutout unless the inspector classifies a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications of the same cutout.

@@

At galaxy-scales, we focus on two primary lens searches. The \rf
\citep{Gavazzi2014} was used for finding compact rings or arcs around centers
of isolated and massive early-type galaxies. By subtracting the PSF-matched
$i$-band images from the $g$-band images, the algorithm looks for excess flux
in the bluer $g$-band. An object detector measures the properties of these
residual blue features, and candidates meeting length-width ratio and
tangential alignment criteria are then visually inspected to form the final
sample. \citet{Gavazzi2014} first selected some 638,000 targets as either
photometrically-classified early type galaxies, or objects selected to have
red centers and blue outer parts, from the T06 CFHTLS data release catalogs. 
14370 were found to show detectable blue residuals, and 2524 were visually
inspected, having passed the automatic feature selection process. This led to
a sample of 42 good quality (\texttt{q\_flag} = 3) and 288 medium quality
(\texttt{q\_flag} = 2) lens candidates. In addition to this well defined
sample, \citet{Gavazzi2014} reported a further 71 serendipitously detected
lens candidates.  From this sample of ``\rf candidates,'' the SL2S team
found, during their follow-up campaign, 39 confirmed lenses (and 17 promising
candidates). We use this sample of 39 ``confirmed \rf lenses'' in our
completeness analysis.

The second galaxy-scale lens search was to find edge-on galaxy lenses in
the \cfhtls \citep{Sygnet2010} by selecting the galaxy's profile from the
output of {\sc Sextractor} in the $i$~band and with low inclination
angle. This sample has about 3 promising and a total of 18 lens
candidates.

On the other hand, the \af \citet{More2012} was used for
finding blue arc-like features in the complete \cfhtls data withouth any
pre-selection on the type of the lensing object. The search was carried
out in the g-band which is the most efficient wavelength to find typical
lensing galaxies. This sample, called the SARCS, has 55 promising and a
total of 127 lens candidates and consists of both galaxy and
groups/cluster scale lens candidates. Arc finding is better suited for
lensed images or arcs with larger image separations i.e. more massive
systems like groups and clusters. Thus, the SARCS sample has mostly
groups/cluster-scale lenses and a few galaxy-scale lenses.

For the purposes of transparency and to help a little with their training, 
the volunteers participating in \sw-\cfhtls lens search were made aware of
these known lens samples. Images containing the systems from the above samples
were labelled as ``Known Lens Candidates'' in the Talk forum, where volunteers
have the opportunity to discuss their findings with other volunteers and the
science team.

@@

{\it State the importance}

@@

We create two main types of simulated lens sample a) galaxy-scale lenses and b)
group or cluster-scale lenses. The galaxy-scale lenses are further divided into two
types based on the nature of the background sources, namely, galaxies and
quasars. Below, we describe how each type of the lens sample was generated.

@@

The $N_{\rm src}$ behind a lens is then calculated by doing the
following integral,

\be
\label{eqn:nsrc}
N_{\rm src} = N_{\rm src}(>L_s,z_s) \int_{z_l}^\infty \sigma_{\rm lens}(\sigma,z_l,z_s,q) D_s^2 (1+z_s)^2 \frac{\rm d\chi}{{\rm d}z_s}
{\rm d}z_s 
\ee

\be
\label{eqn:nlum}
 N_{\rm src}(>L_s,z_s)= \int_{L_{min}}^\infty \Phi(L_s,z_s) {\rm d}L_s
\ee

where $\Phi(L_s,z_s)$ is the source luminosity function per unit comoving
volume, $q$ is the projected axis ratio of the lens ellipticity, $\chi$ and
$D_s$ are the comoving and angular diameter distances to the source,

We use the elliptical galaxy (LRG) catalog from the \cfhtls XXX to select
all the foreground galaxies (e.g. $z<1$) that are potential lenses for the
simulated sample. We exclude all those galaxies whose positions match with the
lensing galaxies from the known \cfhtls~SL2S lens samples \cite{More2012}
within 2~arcsec (XXX check).

First, we calculate the luminosity and velocity dispersion of each potential
lensing galaxy using the CFHT Megacam $g$ and $r$ band magnitudes along with the
photometric redshift ($z_l$) from the LRG catalog.  The Megacam magnitudes are
converted to SDSS magnitudes\footnote{
    http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/megapipe/docs/filters.html} and
are further k-corrected to redshift $z=0.1$ \citep{Frei1994}. We assume that the
evolution of galaxy luminosities is similar to that determined by
\citep{Faber2007}, that is, a decline of $1.5$ in the $m_{r*}$ from redshift $z=1$
to $z=0$ (see Eq.~\ref{magstar}).

\be
\frac{L}{L_{*}}=10.0^{-0.4~(m_{r \rm SDSS}-m_{r*})} 
\ee

where $m_{r*}$ is
\be
\label{magstar}
m_{r*}=-20.44+ 1.5~(z_{l}-0.1) \,.
\ee

We use the $L-\sigma$ relation from
\citep{Parker2005} to get the velocity dispersion as given in Eq.~\ref{magstar2}.

\be
\label{magstar2}
\sigma=142 \left(\frac{L}{L_{*}}\right)^{1/3} 
\ee

Next, in order to decide whether a galaxy is likely to act as a strong lens, we
calculate the lens cross-section ($\sigma_{\rm lens}$) and the number of sources
($N_{\rm src}$) that are in the background. Following \citep{Keeton2000a}, the lens
cross-section is calculated analytically for an isothermal model and is given by

\be
\sigma_{\rm lens}=b_I^2 \, \int_0^{2\pi} 0.5 r^2(\theta) d\theta
\ee

where $b_I$ is 
\be
b_I = b_{\rm SIS} \epsilon_3/sin^{-1}(\epsilon_3) \,,
\ee

the eccentricity ($\epsilon_3$) is 
\be
\epsilon_3=(1-q_3^2)^{1/2}
\ee

and the projected axis ratio is given by
\be
q_k=\sqrt{q_3^2 {\rm sin}^2i_e+{\rm cos}^2i_e} \,.
\ee

In the above equations, $q_3$ is the 3d axis ratio of the ellipsoid and $i_e$ is
the inclination angle. Also, $b_{\rm SIS}= 4\pi
\frac{c^2}{\sigma^2}\frac{D_{s}}{D_{ls}D_l}$ and is referred to as the
Einstein radius where $D_s$,$D_l$ and $D_{ls}$ are angular diameter distances to
the source, the lens and between the lens and source, respectively.

%Comparing \citep{keeton00a} with \citep{kormann94} suggests that $b_I = sqrt(q)*b_{SIS}$

Next, if the foreground galaxy can act as a lens and has at least one source in the
background, then we determine a redshift ($z_s$) and $i$-band magnitude of the
background source(s). We assume two types of background sources namely, galaxies
and quasars. For each source, the redshift and magnitude are generated by drawing randomly from
the following redshift and luminosity distributions. For
galaxies, we assume the redshift distribution is

\be
\label{eqn:ps}
p_s=\frac{\beta z_s^2 {\rm exp}({\frac{z_s}{z_0(m_{\rm lim})}})^\beta}{\Gamma(3/\beta)z_0^3(m_{\rm lim})}
\ee

where $\beta=3/2$ and $z_0(m_{\rm lim})=0.13m_{\rm lim} - 2.2$ and the
luminosity function is

\be
\label{eqn:ns}
n_s=\int^{m_{\rm lim}}_{-\infty} \frac{n_0 {\rm d}m}{\sqrt{10^{2a(m_1-m)}+10^{2b(m_1-m)}}}
\ee

with parameters $a=0.30$, $b=0.56$, $m_1=20$ and $n_0=3\times10^3~deg^{-2}$ as
given in \citep{Faure2009} and references therein. For quasars, we calculate
the luminosity function by following the prescription of \citep{Ogur2010} and
use k-corrections by \citep{Richards2006}.

%alp=-0.5;
%kcorr=-2.5*(1 + alp)*log10(1+zz);
%Dlum=cc.Dlofz(zz)/p.hval;
%DM=5*log10(Dlum);
%Mabs=mag-kcorr-DM-25.0;
The luminosity function is expressed as
\be
\frac{{\rm d}\Phi}{{\rm d}M}=\frac{\Phi_{*}}{10^{0.4(\alpha+1)(M_{\rm abs}-M_{*})} + 10^{0.4(\beta+1)(M_{\rm abs}-M_{*})} }
\ee

where the normalization, $\phi_{*}=5.34\times10^{-6} h^3$ Mpc$^{-3}$ and break
magnitude, $M_*=-20.90 + 5 {\rm log} h - 2.5 {\rm log} f(z)$. The redshift
dependent factor in $M_*$ is given by

\be
f(z)=\frac{e^{\zeta z_s}(1+e^{\xi z_*})}{(\sqrt{e^{\xi z_s}}+\sqrt{e^{\xi z_*}})^2} \,.
\ee
We adopt the best-fit values $\zeta=2.98$, $\xi=4.05$, $z_{*}=1.60$
\citep{Oguri2010}. For the faint end slope, we use $\beta=-1.45$ whereas for
the bright end slope, we use $\alpha=-3.31$ when $z_s<3$ and $\alpha=-2.58$ at
higher redshifts, as prescribed by \citep{Oguri2010}. 
respectively. We note that when calculating $N_{\rm src}$, the source
number density is artificially boosted by a factor (see Table~\ref{tab:thresh})
to increase the occurrence of simulated lenses. This helps in creating a large
enough sample to carry out various performance tests.

>>>>

We report the discovery of 61 new lens candidates in the CFHT Legacy Survey data
(\cfhtls) based on about 11 million classifications performed by citizen
scientists as part of the first blind lens search undertaken by \sw.  The
principle goal of this lens search was to identify lenses missed by lens finding
robots (the \rf on galaxy scales and \af on group/cluster scales), which have
been previously used to mine the \cfhtls data for lenses.  We compare the
properties of known lens samples detected by these algorithms to the \sw sample
and found that they are broadly similar.  The image separation distribution
calculated from the combined sample of known and new lens candidates shows that
our previous constraints on the average density profile of the lens population
are robust.  In addition to recovering about 60\% of the known sample, the new
\sw candidates possess a richer variety compared to the lenses found by the two
robots. We find that analyzing only those classifications which are performed by
the more skilled volunteers, \sw can achieve a detection rate of up to 75\%.
Training and calibration of the performance of citizen scientists is crucial for
the success of \sw. We also present the SIMCT pipeline, used for generating a
sample of realistic simulated lensed images in the CFHTLS, and a sample of duds
and false positives used in the training.  Such a training sample has a legacy
value for testing future lens finding algorithms. We make our training sample
publicly available.

%In summary, \sw
%has been successful in discovering a large sample of new lens candidates in
%addition to recovering a substantial fraction of the known lenses in this pilot
%lens search.

\end{abstract}
%The simulated sample is also used to make a comparison between human performance against the robots.

@@

The last few decades have seen a rise in the discoveries of strong
gravitational lenses owing to the plethora of interesting applications
lenses have in astrophysics and cosmology. Strong lenses are routinely
used to probe the dark matter distribution from galaxy (REF) to cluster
scales (REF), to  study distant young galaxies by using the lensing
magnification as a natural telescope (REF), to test the cosmological
model by constraining cosmological parameters such as the Hubble
constant (REF) and dark energy (REF) and many more. Strong
lenses are rare since a foreground massive object needs to be
sufficiently aligned with a distant background source to produce
multiple images. Nevertheless, systematic lens searches have led to the
discovery of over 500 lenses till
date\footnote{http://masterlens.astro.utah.edu}.

The search for gravitational lenses is a needle-in-a-haystack problem.
Efficient automated methods are imperative to find a reasonably complete and
pure sample of strong lenses. Several lens finding algorithms have been
developed so far for this purpose (REF) but they can not simultaneously capture
the myriad types of lenses that are known to exist.  For example, the lensed
images of background galaxies show variety in their surface brightness
distributions, colors, light profiles, shapes, structures and angular image
separations.  Moreover, many lensed images appear similar to features found
commonly in galaxies (such as spiral arms) or to artifacts in astronomical
images (scattered light around stars).  Almost all lens finding algorithms find
it difficult to distinguish these from the real lenses and thus, suffer from a
high rate of false positives.  To minimize this problem, algorithms are often
restricted to detect a very narrow class of lens systems. However, even after
such restrictions, robotic lens searches have to always rely on visual screening
to arrive at a list of plausible candidates.

Recognising patterns amidst an array of noisy pixelated images is one of the
strengths of human brains (REF). Humans are also capable of dealing with
multi-tier complex web of questions before arriving at a conclusion which may
not be always possible to automate. The algorithm by which our brains process a
task is extremely malleable, self-learning and self-evolving. Therefore it has
a huge potential for the discovery of exotic objects which do not quite fit a
set criteria, but are still very likely to be objects of interest. The lens
finding algorithms are not yet advanced enough to produce better performance
than visual classifications. Given this, participation of a large community of
volunteers to help with the visual identification of lenses is extremely
beneficial for the lensing community. In particular, as we enter the era of
large area imaging surveys (spanning thousands of square degrees), it is
the perfect time to tap into the potential of citizen science.

Galaxy Zoo, was the first citizen science project in astronomy which was used to
classify galaxy morphologies \citep{Lintott2008}.  From the early results based
on morphological classifications of galaxies, to several new unexpected and
interesting discoveries, such as that of green pea galaxies and Hanny's
Voorwerp, Galaxy Zoo has been able to harness the potential of citizen
scientists.  Since then astronomy and non-astronomy projects have been launched
under the citizen science web portal Zooniverse (http://zooniverse.org). The task of finding
gravitational lenses is significantly challenging given the fact that these
lenses are rare. To add to the challenge, not many citizen scientists are
expected to be aware of the phenomenon of gravitational lensing, and the
resulting image configurations.  With these significant challenges at hand, we
designed the \sw project to enable the discovery of lenses through citizen
science.  In Paper I (REF), we describe the design of \sw and how the entire
system functions as a discovery engine.

%
%\subsection{Paper I: Overview}
%In Paper I, we give details of the classification interface used by the
%volunteers. The regular images called test subjects are served randomly from the
%database. Some of the test subjects are labelled as training subjects because
%either these have simulated lensed images superposed around real galaxies or
%there are some false positives in the subjects.  The sample of training subjects
%is much smaller compared to the sample of test subjects. The training subjects
%are shown with decreasing frequency to the volunteers depending on how new the
%volunteers are to the project based on their classification history. However,
%the frequency of training subjects becomes a constant after XXX classifications
%from a volunteer.
%
%Clicking (or failing to click) on the training subjects results in a
%pop-up feedback message that congratulates the volunteers or points them
%to read the Spotter's guide pages. The website has several examples of
%real lenses and false positives with explanation on the Spotter's guide
%page. A quick reference to the Spotter's guide is available in the
%classification interface as well. Additional features to change zoom,
%contrast and brightness are made available to make a better judgement of
%the difficult lens candidates.
%
%The classification of subjects happens in two serial stages. Stage I is
%for quick rejection of all uninteresting subjects. Subjects that passed
%Stage I are then reinspected at Stage II, for a careful selection of
%most promising lens candidates. In general, subjects are assigned a
%small prior probability for it to contain a lens. As each subject is
%seen by multiple volunteers, the posterior probability P is updated by
%using the weights of the volunteers. The weights are determined based on
%how well the volunteers perform on the training sample. Once a subject
%crosses a threshold (acceptance or rejection), it is retired from the
%database to ensure that rest of the subjects receive sufficient multiple
%classifications. This is true for stage I with ~10$^5$ subjects but is
%not required for stage II with only a few thousand subjects.
%
%{\bf AM: add one more paragraph that gives some key conclusions and numbers
%from Paper I}
%\question{AM}{PJM,AV, should we have this overview section on paper I?}
%

In this paper ( Paper II hereafter), we describe our first lens search using
data from the Canada-France-Hawaii Telescope Legacy Survey (\cfhtls, REF).  This
paper is organised as follows. In \Sref{sec:data}, we introduce the \cfhtls
imaging data and the known lens samples from the \cfhtls. In \Sref{sec:ts} and
\Sref{sec:dfp}, we explain the training sample consisting of simulated lenses,
duds and false positives generated to aid the \sw volunteers in the lens search.
In \Sref{sec:results}, we present the new lens candidates from \sw and compare
\sw to two lens finding algorithms which produced the known \cfhtls lens
samples. Next, we discuss what kind of lenses are detected or missed by the
algorithms and \sw in \Sref{sec:discuss}. The summary and conclusions are given
in \Sref{sec:conclude}.

@@

The \cfhtls is a photometric survey in five optical bands
($u^*g'r'i'z'$) carried out with the wide-field imager MegaPrime which
has a 1~deg$^2$ field-of-view and a pixel size of 0.186\arcsec. The
\cfhtls \textsc{WIDE} covers a total area of 171~deg$^2$ on the sky and
consists of four fields W1, W2, W3 and W4. The field W1 has the
largest sky coverage of 63.65~deg$^2$. The fields W2 and W4 have similar
sky coverages of 20.32~deg$^2$ and 20.02~deg$^2$,
respectively\footnote{These numbers are estimated from
http://terapix.iap.fr/cplt/table\_syn\_T0006.html}.  The field W3 has a
sky coverage of 42.87~deg$^2$ and is more than twice as large as W2 and
W4.

@@

We note that the \cfhtls is a niche survey with a unique combination of wide
imaging with deep sensitivity. It is a precursor to the ongoing wide imaging
surveys such as the Dark Energy Survey (DES), Kilo Degree Survey (KiDS) and the
Hyper Suprime-Cam (HSC) survey and other planned future surveys such as the
Large Synoptic Survey Telescope (LSST) survey.  The search for lenses with \sw
in the \cfhtls is an important step to learn lessons and prepare for lens
searched in these larger imaging surveys.

@@

In order to perform a blind lens search over the entire \cfhtls \textsc{wide},
we present the users with cutouts of images selected from the survey region. We
use the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for visual
identification of lenses. We also provided access to the $u$ and $z$ band imaging
data through dashboard, when users wished to take a closer look at a particular system\footnote{The dashboard tool page can be accessed from
http://tools.zooniverse.org/\#/dashboards/spacewarp}.

We made color composite images using the publicly-available
code HumVI\footnote{The open source  color image composition code used in this
work is available from \texttt{http://github.com/drphilmarshall/HumVI}}
following the prescription of \citet{Lupton2004}. The code
rescales the pixel values of each channel image into flux units, and
then applies an arcsinh stretch. The stretch parameters are chosen
using a small random sample of images, to ensure that the background
noise is just visible, and that the centres of bright, intermediate
redshift galaxies are not saturated. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to
allow straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) with an
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This  results in a catalog of $\sim$ 430 000 cutouts for the entire \cfhtls
\textsc{wide} region. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. It is possible that a lens
candidate happens to be too close to the edge of a cutout, the overlap
between neighbouring cutout allows the inspector to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, an inspector may not necessarily come
across the neighbouring cutout unless the inspector classifies a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.

@@

At galaxy-scales, two primary lens searches have been carried out in
\cfhtls. The \rf \citep{Gavazzi2014} was used for finding compact rings
or arcs around centers of isolated and massive early-type galaxies. The
\rf subtracts the PSF-matched $i$-band images from the $g$-band images,
and looks for excess flux in the bluer $g$-band. An object detector
measures the properties of these residual blue features, and candidates
which meet the length-width ratio and tangential alignment criteria are
then visually inspected to form the final sample. \citet{Gavazzi2014}
pre-selected $\sim$638,000 targets as either photometrically-classified
early type galaxies, or objects selected to have red centers and blue
outer parts, from the T06 CFHTLS data release catalogs.  14370 were
found to show detectable blue residuals, and 2524 were visually
inspected, having passed the automatic feature selection process. This
led to a total of 330 lens candidates out of which 42 are good quality
(\texttt{q\_flag} = 3) and 288 are medium quality (\texttt{q\_flag} = 2)
candidates. In addition to this well defined sample, \citet{Gavazzi2014}
reported a further 71 serendipitously detected lens candidates.  From
this sample of ``\rf candidates,'' the SL2S team found, during their
follow-up campaign, 33 confirmed lenses (and 17 promising candidates).
{\bf AM: check if any of these are from \cfhtls DEEP?}

The second galaxy-scale lens search was to find edge-on galaxy lenses in the
\cfhtls \citep{Sygnet2010} by selecting galaxies based on profile fits in the
$i$-band performed using {\sc Sextractor} and with low inclination angle. This
sample has about 3 promising and a total of 18 lens candidates.
{\bf AM: perhaps, this should go in additional samples because this was not
used as one of the known samples,check ?}

Lastly, the \af \citep{More2012} was used for finding blue arc-like features in
the complete \cfhtls data without any pre-selection on the type of the lensing
object. The algorithm measures the second order moments of the flux distribution
in pixels within small regions to estimate the direction and extent of local
elongation of features. Pixels with high values of elongation are connected to
form an arc candidate. Finally, a set of thresholds on arc properties such as
the area, length, width, curvature and surface brightness are used to select
arc-like candidates. The search was carried out in the g-band which is the most
efficient wavelength to find typical lensed features.  This sample, called
SARCS, has 55 promising and a total of 127 lens candidates which are selected
from both \cfhtls WIDE and DEEP fields. The SARCS sample consists of some
galaxy-scale candidates and mostly groups/cluster scale lens candidates. This is
because more massive systems produce arcs or lensed images with large image
separation from the lensing galaxy which are easier to detect compared to the
galaxy-scales.  In the absence of a large systematically followed up
verification of the candidates, we use the  most promising sample of 26 lens
candidates as the bonafide known lens sample. And, the sample of 108 candidates
form the known lens candidates which come from the \cfhtls WIDE alone since the
\sw lens search is carried out on the WIDE imaging only.

For the purposes of transparency and to help with the training, 
the volunteers participating in \sw-\cfhtls lens search were made aware of
these known lens samples. Images containing the systems from the above samples
were labelled as ``Known Lens Candidates'' in the Talk forum, where volunteers
have the opportunity to discuss their findings with fellow volunteers and the
science team. In this paper, we refer to the sample of 330 \rf and 108
\af lens candidates as the sample of ``known lens candidates'' and the sample of
confirmed (or most promising) 33 \rf and 26 \af as the sample of
``known lenses''.

\question{AM}{AV, add info on other lens samples from the literature?}

@@

In this section, we describe the framework to generate the simulated lens sample
and give details of the sample itself along with some known limitations. This
simulated lens sample is an important component for analyzing the
classifications performed by the citizen scientists.

@@

For the purpose of generating simulated gravitational systems, we divide the
lenses in to two main categories a) galaxy scale lenses b) and group or
cluster-scale lenses. We further subdivide galaxy scale lenses based on the
type of the background sources, namely galaxies and quasars. We do not simulate
group scale quasar lenses as they are expected to be extremely rare. We now
describe our procedure to generate these different types of lens systems.

@@

We begin by considering all elliptical galaxies at $z<1$ in our parent \cfhtls
catalog as potential lens candidates for the simulated sample. To avoid using a
known lens galaxy for our simulation purpose, we exclude all those galaxies
whose positions match with the lensing galaxies from the known \cfhtls~SL2S lens
samples \cite{More2012} within 2~arcsec (XXX check).

For each galaxy, the average number of source objects (either quasars or
galaxies) above a minimum luminosity $L_{\rm min}$ in the background that may get lensed
can be calculated as
\be
\label{eqn:nsrc}
N_{\rm src} = \int_{z_l}^\infty n_{\rm src}(>L_{\rm min},z_s)  \sigma_{\rm
lens}(\sigma_v,z_l,z_s,q) \frac{\rm d V}{{\rm d}z_s}
{\rm d}z_s 
\ee
where
\be
\label{eqn:nlum}
n_{\rm src}(>L_{min},z_s)= \int_{L_{\rm min}}^\infty \Phi(L',z_s) {\rm d}L'
\ee
Here, $\Phi(L',z_s)$ denotes the source luminosity function per unit comoving
volume, $\sigma_{\rm lens}$ denotes the angular lens cross-section, which depends upon
the lens redshift ($z_l$), source redshift ($z_s$), the lens velocity
dispersion $\sigma_v$ as well as the projected axis ratio of the lens
ellipticity, $q$.

In order to calculate the lensing cross-section, we first calculate the
luminosity of each potential lensing galaxy using the photometric redshifts
($z_l$) from the LRG catalog. We use the $L-\sigma$ scaling relation from the
bright sample of \citep{Parker2007} to set the velocity dispersion of the halo
hosting the galaxy which will be later used in the model
\be
\label{magstar2}
\sigma_v=142 \left(\frac{L}{L_{*}}\right)^{1/3} \,.
\ee
We assume that the knee of the luminosity function of galaxies, $L_*$ evolves
such that there is a decline of $1.5$ magnitudes between $z=1$ to $z=0$
\citep{Faber2007}.\footnote{We anchor our $L_*$ evolution at low redshifts using the
determination of $L_*$ in the r-band by \citet{Blanton:2001}. To maintain
consistency in magnitude systems, we have converted the CFHT MegaCAM magnitudes
to SDSS magnitudes and k-corrected them to $z=0.1$.}

%first calculate the
%luminosity and velocity dispersion of each potential lensing galaxy using the
%CFHT Megacam $g$ and $r$ band magnitudes along with the photometric redshift
%($z_l$) from the LRG catalog.  The Megacam magnitudes are converted to SDSS
%magnitudes\footnote{
%http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/megapipe/docs/filters.html} and are
%further k-corrected to redshift $z=0.1$ \citep{Frei1994}. We assume that the
%evolution of galaxy luminosities is similar to that determined by
%\citep{Faber2007}, that is, a decline of $1.5$ in the $m_{r*}$ from redshift
%$z=1$ to $z=0$ (see Eq.~\ref{magstar}).

%\be
%\frac{L}{L_{*}}=10.0^{-0.4~(m_{r \rm SDSS}-m_{r*})} 
%\ee
%
%where $m_{r*}$ is
%\be
%\label{magstar}
%m_{r*}=-20.44+ 1.5~(z_{l}-0.1) \,.
%\ee
%
%We use the $L-\sigma$ relation from
%\citep{Parker2005} to get the velocity dispersion as given in Eq.~\ref{magstar2}.

%\be
%\label{magstar2}
%\sigma=142 \left(\frac{L}{L_{*}}\right)^{1/3} 
%\ee

We assume a singular isothermal ellipsoid model for each of our galaxies
\citep{Kormann1994}, such that the convergence is given by
\be
\kappa (x,y) = \frac{b \sqrt{q}}{2}\frac{1}{\left( \theta_1^2 + q^2\theta_2^2\right)^2}
\ee
Here b is called the Einstein radius, and its dependence on the velocity
dispersion of the SIE is given by
\be
b = 4\pi\,
\left(\frac{\sigma_v^2}{c^2}\right)\left(\frac{D_{ls}}{D_{s}}\right) \,.
\ee
The SIE model results in a caustic and a pseudo-caustic on the source plane:
which demarcate the regions of different image multiplicities. Parametric
solutions, $r(\theta)$ (where $\theta$ is the polar angle), for the caustics in
such a model were given by \citet{Keeton2000b}. We use these parametric
solutions to obtain the area of the lensing cross-section, $\sigma$, for every
galaxy,
\be
\sigma_{\rm lens}=\frac{b^2 q}{2} \, \int_0^{2\pi} r^2(\theta) d\theta\,.
\ee
At every polar angle we use the maximum of the radial and tangential caustic for
the purpose of calculating the cross-section.

%where
%\be
%b_I =  \sqrt{q}\, 4\pi\,
%\left(\frac{\sigma_v^2}{c^2}\right)\left(\frac{D_{ls}}{D_{s}}\right)\,.
%\ee
%Here $D_s$ and $D_{ls}$ are angular diameter distances to
%the source, and between the lens and source, respectively.

%Next, if the foreground galaxy can act as a lens and has at least one source in the
%background, then we determine a redshift ($z_s$) and $i$-band magnitude of the
%background source(s). We assume two types of background sources namely, galaxies
%and quasars. For each source, the redshift and magnitude are generated by drawing randomly from
%the following redshift and luminosity distributions. For
%galaxies, we assume the redshift distribution is

We use the results of \citet{Faure2009} to specify the source luminosity
function of galaxies. We assume that the redshift distribution of sources is
given by
\be
\label{eqn:ps}
p_s=\frac{\beta z_s^2 {\rm exp}({\frac{z_s}{z_0(m_{\rm lim})}})^\beta}{\Gamma(3/\beta)z_0^3(m_{\rm lim})}
\ee
where $\beta=3/2$ and $z_0(m_{\rm lim})=0.13m_{\rm lim} - 2.2$ and the source
counts as a function of the limiting magnitude are given by
\be
\label{eqn:ns}
n_s=\int^{m_{\rm lim}}_{-\infty} \frac{n_0 {\rm d}m}{\sqrt{10^{2a(m_1-m)}+10^{2b(m_1-m)}}}
\,,
\ee
with parameters $a=0.30$, $b=0.56$, $m_1=20$ and $n_0=3\times10^3~deg^{-2}$.
